Automated Formal Synthesis of Digital Controllers for State-Space Physical Plants

### 2017-01-24 cav2017/cav2017.tex

Linear Time Invariant (LTI) control models represent a broad class of dynamical systems
%modern embedded software, performing highly non-trivial control tasks,
with significant impact in numerous application areas such as
the life sciences, robotics, and engineering~\cite{astrom1997computer,Franklin15}.
Control synthesis for LTI is broadly investigated, 
however the use of digital control architectures adds new challenges due to 
%exacerbated by specific artefacts,  
the effects of finite-precision arithmetic, time discretization, and quantization noise, which are typically introduced 
by Analogue-to-Digital (A/D) and Digital-to-Analogue (D/A) conversion.  
Whilst the research on digital control is well developed~\cite{astrom1997computer}, 
automated and sound control synthesis is a challenging task. 
%Thus, the programming is a key barrier to broad adoption of digital control, 
%and requires considerable expertise.

Although there is a vast literature on validation of control systems,
the subject of control synthesis has been somewhat restricted in the
properties it explores.  Traditionally, this focuses mainly on stability %, working typically on continuous domains
\cite{DBLP:journals/corr/AbateBCCDKK16,sadabadi2016static}, or using switched
models~\cite{DBLP:conf/emsoft/RavanbakhshS16}.
Reachability tools are
also used when discretizing the state-space for verification and
parametric synthesis~\cite{cimatti2013parameter}.

In this work, we propose the synthesis of \emph{safe} control algorithms for
LTI models taking into consideration the continuous domain of the plant, 
alongside the discrete domain of the controller and the hybrid elements that relate them.  
Due to the complexity of such systems, in this particular work we focus
on linear models with known  implementation features (e.g., number of bits, fixed-point
arithmetic) and perform synthesis of safe digital controllers.

We describe and evaluate two approaches for synthesizing digital controllers 
for state-space physical plants. Both approaches consider a counterexample 
guided inductive synthesis (CEGIS) loop. In the first approach, 
we devise a digital controller that stabilizes the system whilst remaining safe given a 
pre-selected time horizon; then, we verify its 
potential safety by unfolding the dynamics of the system and checking for 
a completeness threshold. In the second approach, we employ abstract acceleration
to accelerate the dynamics forward instead of symbolically unwinding the 
transition relation~\cite{cattaruzza2015unbounded}. We provide experimental results 
showing that both approaches are able to efficiently synthesize stable and safe controllers 
for a set of intricate physical plant models taken from the digital 
control literature: the median runtime for our benchmark
set is \textcolor{red}{$XXX$}s, i.e., most controllers can be synthesized in less
than \textcolor{red}{$XXX$} seconds.

@@

In this study, we are not limited to the synthesis of digital stabilizing controllers -- 
a well known task in the literature on digital control system -- 
but target safety requirements with an overall approach that is sound and automated. 
More specifically, 
we require that the closed-loop system \eqref{eq:closedloopss} meets given safety specifications. 
A safety specification raises a requirement on the states of the model, 
such that the feedback controller (namely the choice of the gains in $K$) 
must ensure that the states never exist such requirement. 
Note that a stable, closed-loop system is not necessarily a safe system: 
indeed, the state values may leave the safe whilst they converge to the equilibrium, 
which is typical in the case of oscillatory dynamics. However, a safe system will always be stable.
In this work, the safety property represented by $\phi_{safety}$ is encoded as follows: 
%
\begin{equation}
\label{eq:safetyliteral}
\phi_{safety}\iff \forall k\geq 0, \bigwedge_{i=1}^{n}{\underline{x_{i}} \leq x_{i,k} \leq \overline{x_{i}}},
\end{equation}
%
%\addtodo{We consider $x_{i}^{-}$ and $x_{i}^{+}$ to be -1 and 1, respectively.}
where $\underline{x_{i}}$ and $\overline{x_{i}}$ are lower and upper bounds for the $i$-th coordinate $x_{i}$ of state $x\in \mathbb R^n$ at the $k$-th instant, respectively. 
This means that the states will always be within the $n$-dimensional hyper-box. 
%, which is typically defined by the safety specification.

>>>>

Linear Time Invariant (LTI) control models represent a broad class of dynamical systems
%modern embedded software, performing highly non-trivial control tasks,
with significant impact in numerous application areas as in 
the life sciences, robotics, and engineering~\cite{astrom1997computer,Franklin15}.
Control synthesis for LTI is broadly investigated, 
however the use of digital control architectures adds new challenges due to 
%exacerbated by specific artefacts,  
the effects of finite-precision arithmetic, time discretization, and quantization noise, which are typically introduced 
by Analogue-to-Digital (A/D) \todo{later ADC and DAC} and Digital-to-Analogue (D/A) conversion.  
Whilst the research on digital control is well developed~\cite{astrom1997computer}, 
automated and sound control synthesis is a challenging task,  
%Thus, the programming is a key barrier to broad adoption of digital control, 
%and requires considerable expertise.
%Whilst there is a vast literature on validation of control systems,
%the subject of control synthesis has been somewhat restricted in the properties it explores.  
%Traditionally, this focuses mainly on stability %, working typically on continuous domains
particularly when the synthesis objective goes beyond classical stability or tracking requirements. 
%\cite{DBLP:journals/corr/AbateBCCDKK16,sadabadi2016static}, or using switched
%models~\cite{DBLP:conf/emsoft/RavanbakhshS16}. 
Beyond stability, 
recently reachability tools have been developed [cite SPACEx, Phaver, and others], 
however they have not been employed towards control synthesis. 
\todo{why do we cite this paper \cite{cimatti2013parameter}?what is the relationship with our work? }
% when discretising the state-space for verification and
%parameter synthesis~\cite{cimatti2013parameter}.

In this work, we propose the synthesis of \emph{safe} control algorithms for
LTI models taking into consideration the continuous domain of the plant, 
alongside the discrete domain of the controller, 
as well as the hybrid elements that relate them.  
%% In this paper, we synthesize controllers for hybrid
%% closed-loop systems, i.e., software-implemented embedded controllers along
%% with a model of their physical environment (the plant), that are
%% \emph{safe}.
Due to the complexity of such systems, in this work we focus 
on linear models with known  implementation features (e.g., number of bits, fixed-point
arithmetic), %\addtodo{[clarify next] with known configurations} 
and perform %parametric 
synthesis of safe digital controllers - the extension to more general quantitative properties is left to future work.

We describe and evaluate two approaches for synthesising digital controllers 
for state-space physical plants. Both approaches consider a counterexample 
guided inductive synthesis (CEGIS) loop. In the first approach, 
we devise a digital controller that stabilizes the system whilst remaining safe given a 
pre-selected time horizon; then, we verify its 
potential safety by unfolding the dynamics of the system and checking for 
a completeness threshold. In the second approach, we substitute the second step by employing abstract acceleration
to unfold the dynamics forward, instead of symbolically unwinding the 
transition relation~\cite{cattaruzza2015unbounded}. 
\todo{We need to mention the CEGIS architecture here. We need to emphasise the derivation of errors for soundness. Can we claim completeness already?}
We provide experimental results 
showing that both approaches are able to efficiently synthesise stable and safe controllers 
for a set of intricate physical plant models taken from the digital 
control literature: the median runtime for our benchmark
set is \textcolor{red}{$XXX$}s, i.e., most controllers can be synthesized in less
than \textcolor{red}{$XXX$} seconds.

@@

In this study, we are not bound to the synthesis of digital stabilizing controllers -- 
a well known task in the literature on digital control system -- 
but target safety requirements with an overall approach that is sound and automated. 
More specifically, 
we require that the closed-loop system \eqref{eq:closedloopss} meets given safety specifications. 
A safety specification raises a requirement on the states of the model, 
such that the feedback controller (namely the choice of the gains in $K$) 
must ensure that the states never exist such requirement. 
Note that a stable, closed-loop system is not necessarily a safe system: 
indeed, the state values may leave the safe whilst they converge to the equilibrium, 
which is typical in the case of oscillatory dynamics. However, a safe system will always be stable.
In this work, the safety property represented by $\phi_{safety}$ is encoded as follows: 
%
\begin{equation}
\label{eq:safetyliteral}
\phi_{safety}\iff \forall k\geq 0, \bigwedge_{i=1}^{n}{\underline{x_{i}} \leq x_{i,k} \leq \overline{x_{i}}},
\end{equation}
%
%\addtodo{We consider $x_{i}^{-}$ and $x_{i}^{+}$ to be -1 and 1, respectively.}
where $\underline{x_{i}}$ and $\overline{x_{i}}$ are lower and upper bounds for the $i$-th coordinate $x_{i}$ of state $x\in \mathbb R^n$ at the $k$-th instant, respectively. 
This means that the states will always be within the $n$-dimensional hyper-box. 
%, which is typically defined by the safety specification.
